### 2018-06-08 doc/main.tex
Added source for documentation

\documentclass[a4paper,parskip=never,UKenglish]{scrartcl}
\input{preamble}
\title{Rusty Variation}
\subtitle{Affine Session Types for Rust}
\author{%
  Wen Kokke\protect\\
  Supervised by Philip Wadler and Ian Stark\protect\\
  Graduate student (8450698)}
\affil{
  LFCS, University of Edinburgh\protect\\
  Informatics Forum, 10 Crichton St, Edinburgh EH8 9AB, UK\protect\\
  \email{wen.kokke@ed.ac.uk}}
\date{June 16, 2018}
\begin{document}
\maketitle
\begin{abstract}
  Rusty Variation (RV) is a library for session-typed communication in Rust, which offers strong compile-time correctness guarantees. Programs written using RV are guaranteed to adhere to a specified protocol, and are guaranteed to be free from dead- and livelocks. RV is an implementation of Exceptional GV, a session-typed functional language with exceptions and session cancellation.
\end{abstract}

\section{Problem and Motivation}

\section{Background and Related Work}
\subsection{Session Types for Rust}
\subsection{Exceptional GV}

\section{Approach and Uniqueness}

\section{Results and Contributions}
\end{document}